---
title: "Untitled"
author: "Lester"
date: "2022-12-16"
output: html_document
---
```{r}
library(pictograph)
library(tidyverse)
library(rjags)
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/bic.R")
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/cluster_separately.R")
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/clustering_functions.R")
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/plot.R")
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/gabowmyers.R")
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/graph_ops.R")
# source("C:/Users/Leste/OneDrive - Johns Hopkins/Documents/GitHub/pictograph/R/metrics.R")
get.parameter.chain <- function(param, chains) {
  chains[grep(paste0(param, "\\["), chains$Parameter), ]
}
```

#replace y<10 with 0
```{r}
input_data <- readRDS("C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/pictograph-input/IP22_input_data.rds")
ny <- input_data$y
ny[ny<10] = 0
mut_in <- as.data.frame(ny)%>%
  mutate(sum= rowSums(.))
#input_data$y = ny

input_data <- list(y = input_data$y[mut_in$sum!=0,],
              n = input_data$n[mut_in$sum!=0,],
              tcn = input_data$tcn[mut_in$sum!=0,],
              m = input_data$m[mut_in$sum!=0,],
              I = sum(mut_in$sum!=0),
              S = input_data$S,
              purity = input_data$purity,
              Sample_names = input_data$Sample_names,
              Mut_ID = input_data$Mut_ID[mut_in$sum!=0])

# all_set_results <- clusterSep(input_data,
#                              n.iter = 1000, n.burn = 100, thin = 1,
#                              max_K = 5)
# saveRDS(all_set_results,"C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/remove_less10/IP24_input_data_bt10.rds")
all_set_results <- readRDS("C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/remove_less10/IP22_input_data_bt10.rds")

set_k_choices <- writeSetKTable(all_set_results)
set_k_choices$chosen_K[5] = 2


best_set_chains <- collectBestKChains(all_set_results, chosen_K = set_k_choices$chosen_K)
chains <- mergeSetChains(best_set_chains, input_data)
plotChainsCCF(chains$w_chain)


sets <- writeClusterAssignmentsTable(chains$z_chain)%>%
  mutate(Mut_ID = as.numeric(gsub("Mut","",Mut_ID)))%>%
  arrange(Mut_ID)
mut <- input_data$Mut_ID%>%
  cbind(sets)%>%
  rename("Mut" = ".")%>%
  filter(grepl('KRAS|GNAS',Mut, ignore.case = TRUE))



w_mat <- estimateCCFs(chains$w_chain)
graph_G_pre <- prepareGraphForGabowMyers(w_mat, chains, input_data)
graph_G <- filterEdgesBasedOnCCFs(graph_G_pre, w_mat, thresh = 0.1)
enumerateSpanningTreesModified(graph_G, w_mat, sum_filter_thresh=0.59)
scores <- calcTreeScores(chains$w_chain, all_spanning_trees)
best_tree <- all_spanning_trees[[which.max(scores)]]
plotTree(best_tree)
plotEnsembleTree(all_spanning_trees)
subclone_props <- calcSubcloneProportions(w_mat, best_tree)
subclone_props[subclone_props<0]=0
subclone_props_n <- sapply(c(1,2,3), function(x) round(subclone_props[,x]/sum(subclone_props[,x]),digits = 5))
plotSubclonePie(subclone_props_n)
plotSubcloneBar(subclone_props_n)
```

# hist for y/n
# ```{r setup, include=FALSE}
# input_data <- readRDS("C:/Users/Leste/Desktop/Karchin Lab/data/pictograph-input/IP22_input_data.rds")
# y <- input_data$y
# n <- input_data$n
# # check the distribution of y/n
# #y[y < 10] = NA
# per <- y/n
# his_in <- c(per)
# hist(his_in)
# ```


# get liklihood for the model
# y is predicted from a binomial distritbuion with parameter VAF and n (from model output), vaf can be calc using w chain. 
# then find the liklihood of observed data form the expected distribution
```{r}
input_data <- readRDS("C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/pictograph-input/IP24_input_data.rds")
all_set_results <- readRDS("C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/IP24_processed.rds")
all_types <- names(all_set_results)
AAIC <- list()
for (i in 1:length(all_types)){
  all_k <- names(all_set_results[[i]]$all_chains)
  AIC <- matrix(NA, nrow = 2, ncol = length(all_k))
  col = 1
  for (j in 1:length(all_k)){
    chains <- all_set_results[[i]]$all_chains[[j]]
    w <- chains$w_chain %>%
      mutate(Cluster = as.numeric(gsub("w\\[(.*),.*","\\1",Parameter)),
             sample = gsub(".*,(.*)\\]","\\1", Parameter),
             value = round(value,1))%>%
      group_by(Cluster, sample)%>%
      summarize(w = mean(value))%>%
      ungroup()%>%
      spread(key = sample, value = w)%>%
      arrange(Cluster)
    
    ww <- writeClusterAssignmentsTable(chains$z_chain)
      mutate(Mut_ID = as.numeric(gsub("Mut","",Mut_ID)))%>%
      arrange(Mut_ID)%>%
      left_join(w)%>%
      select(-c("Mut_ID","Cluster"))%>%
      as.matrix()
    
    pres <- input_data$y %>%
      as.data.frame()%>%
      mutate(across(everything(), ~ifelse(.x >0, 1, 0)))%>%
      unite("type", colnames(.), sep= "", remove = FALSE)
    
    rows <- pres$type == all_types[i]
    
    y <- input_data$y[rows,]
    n <- input_data$n[rows,]
    m <- input_data$m[rows,]
    p <- input_data$purity
    pp <- diag(p,length(p),length(p))
    c <- input_data$tcn[rows,]
    
    vaf <- m*ww%*%pp/sweep(c%*%pp, 2, -2*(1-p))
    
    # vaf is the probability
    # trails is n
    #assume predicted n is from the input
    lik <- prod(dbinom(y,n,vaf))
    # 5 parameters? 
    aic_score = 2*-2*log(lik)
    
    AIC[1,col] = all_k[j]
    AIC[2,col] = aic_score
    
    col=col+1
  }
  AAIC[[all_types[i]]] <- AIC
}

aic_k <- gsub("K","",sapply(1:length(AAIC), function(x) AAIC[[x]][1,which.min(AAIC[[x]][2,])]))


## AIC = 2K – 2ln(L)

```


# cluster.sample.presence[[j]] : subscript out of bounds: means some cluster doesn't exist, reduce chosen k

```{r}
set_k_choices <- writeSetKTable(all_set_results)%>%
  mutate(chosen_K = as.numeric(aic_k))
set_k_choices$chosen_K[4] = 1

best_set_chains <- collectBestKChains(all_set_results, chosen_K = set_k_choices$chosen_K)
chains <- mergeSetChains(best_set_chains, input_data)
plotChainsCCF(chains$w_chain)


## check mutation location on the tree ##
sets <- writeClusterAssignmentsTable(chains$z_chain)%>%
  mutate(Mut_ID = as.numeric(gsub("Mut","",Mut_ID)))%>%
  arrange(Mut_ID)
mut <- input_data$Mut_ID%>%
  cbind(sets)%>%
  rename("Mut" = ".")%>%
  filter(grepl('KRAS|GNAS',Mut, ignore.case = TRUE))

```


#tree
```{r}
w_mat <- estimateCCFs(chains$w_chain)
graph_G_pre <- prepareGraphForGabowMyers(w_mat, chains, input_data)
graph_G <- filterEdgesBasedOnCCFs(graph_G_pre, w_mat, thresh = 0.1)
enumerateSpanningTreesModified(graph_G, w_mat, sum_filter_thresh=0.5)
scores <- calcTreeScores(chains$w_chain, all_spanning_trees)
best_tree <- all_spanning_trees[[which.max(scores)]]
best_tree <- all_spanning_trees[[10]]
plotTree(best_tree)
plotEnsembleTree(all_spanning_trees)
subclone_props <- calcSubcloneProportions(w_mat, best_tree)
subclone_props[subclone_props<0]=0
subclone_props_n <- sapply(c(1,2,3), function(x) round(subclone_props[,x]/sum(subclone_props[,x]),digits = 5))
plotSubclonePie(subclone_props_n)
plotSubcloneBar(subclone_props_n)
```