---
title: "Untitled"
author: "Lester"
date: "2022-12-16"
output: html_document
---
```{r}
library(pictograph)
library(tidyverse)
library(rjags)

```

#replace y<10 with 0
```{r}
input_data <- readRDS("C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/pictograph-input/IP22_input_data.rds")
ny <- input_data$y
ny[ny<10] = 0

#check if the sample presence is 000
mut_in <- as.data.frame(ny)%>%
  mutate(sum= rowSums(.))
#input_data$y = ny

input_data <- list(y = input_data$y[mut_in$sum!=0,],
              n = input_data$n[mut_in$sum!=0,],
              tcn = input_data$tcn[mut_in$sum!=0,],
              m = input_data$m[mut_in$sum!=0,],
              I = sum(mut_in$sum!=0),
              S = input_data$S,
              purity = input_data$purity,
              Sample_names = input_data$Sample_names,
              Mut_ID = input_data$Mut_ID[mut_in$sum!=0])

# all_set_results <- clusterSep(input_data,
#                              n.iter = 1000, n.burn = 100, thin = 1,
#                              max_K = 5)
# saveRDS(all_set_results,"C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/remove_less10/IP24_input_data_bt10.rds")
all_set_results <- readRDS("C:/Users/Leste/OneDrive - Johns Hopkins/Desktop/Karchin Lab/data/remove_less10/IP22_input_data_bt10.rds")

set_k_choices <- writeSetKTable(all_set_results)
set_k_choices$chosen_K[5] = 2


best_set_chains <- collectBestKChains(all_set_results, chosen_K = set_k_choices$chosen_K)
chains <- mergeSetChains(best_set_chains, input_data)
plotChainsCCF(chains$w_chain)


sets <- writeClusterAssignmentsTable(chains$z_chain)%>%
  mutate(Mut_ID = as.numeric(gsub("Mut","",Mut_ID)))%>%
  arrange(Mut_ID)
mut <- input_data$Mut_ID%>%
  cbind(sets)%>%
  rename("Mut" = ".")%>%
  filter(grepl('KRAS|GNAS',Mut, ignore.case = TRUE))



w_mat <- estimateCCFs(chains$w_chain)
graph_G_pre <- prepareGraphForGabowMyers(w_mat, chains, input_data)
graph_G <- filterEdgesBasedOnCCFs(graph_G_pre, w_mat, thresh = 0.1)
enumerateSpanningTreesModified(graph_G, w_mat, sum_filter_thresh=0.59)
scores <- calcTreeScores(chains$w_chain, all_spanning_trees)
best_tree <- all_spanning_trees[[which.max(scores)]]
plotTree(best_tree)
plotEnsembleTree(all_spanning_trees)
subclone_props <- calcSubcloneProportions(w_mat, best_tree)
subclone_props[subclone_props<0]=0
subclone_props_n <- sapply(c(1,2,3), function(x) round(subclone_props[,x]/sum(subclone_props[,x]),digits = 5))
plotSubclonePie(subclone_props_n)
plotSubcloneBar(subclone_props_n)
```

# hist for y/n
```{r}
input_data <- readRDS("C:/Users/Leste/Desktop/Karchin Lab/data/pictograph-input/IP22_input_data.rds")
y <- input_data$y
n <- input_data$n
# check the distribution of y/n
y[y < 10] = NA
per <- y/n
his_in <- c(per)
hist(his_in)
```